---
title: "sub_suduko"
author: "Tynan"
date: "2023-10-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(magrittr)
```

## R Markdown

You are working on a logic game made up of a series of puzzles. 

The first type of puzzle you settle on is "sub-Sudoku", a game where the player has to position the numbers 1..N on an NxN matrix.

Your job is to write a function that, given an NxN matrix, returns true if every row and column contains the numbers 1..N

The UI for the game does not do any validation on the numbers the player enters, so the matrix can contain any signed integer.

```{r test_data_raw}
# Examples:

# [[1, 2, 3],
#  [3, 1, 2],
#  [2, 3, 1]]        -> True

# [[1, 2, 3],
#  [1, 2, 3],
#  [1, 2, 3]]        -> False

# [[1, 1, 1],
#  [2, 2, 2],
#  [3, 3, 3]]        -> False

# [[1000, -1000, 6],
#  [   2,     3, 1],
#  [   3,     1, 2]] -> False

# [[0]]              -> False

# [[3, 2, 3, 2],
#  [2, 3, 2, 3],
#  [3, 2, 3, 2],
#  [2, 3, 2, 3]]     -> False

# [[2, 3, 4],
#  [3, 4, 2],
#  [4, 2, 3]]        -> False

# [[-1,-2,-3],
#  [-2,-3,-1], 
#  [-3,-1,-2]]       -> False

# [[1,1,1],
#  [1,1,2],
#  [1,2,3]]          -> False

# [[1]]              -> True

# n: The number of rows/columns in the matrix

```

```{r test_data}
# True
grid1 <- matrix(c(1,2,3,
	                3,1,2,
	                2,3,1), nrow=3, ncol=3, byrow=TRUE)
grid1                  

# False
grid2 <- matrix(c(1,2,3,
	                1,2,3,
	                1,2,3), nrow=3, ncol=3, byrow=TRUE)
# False
grid3 <- matrix(c(1,1,1,
	                2,2,2,
	                3,3,3), nrow=3, ncol=3, byrow=TRUE)
	                
# False
grid4 <- matrix(c(1000,-1000,6,
	                2,3,1,
	                3,1,2), nrow=3, ncol=3, byrow=TRUE)
	                
# False
grid5 <- matrix(c(0), nrow=1, ncol=1, byrow=TRUE)

# False
grid6 <- matrix(c(3, 2, 3, 2,
	                2, 3, 2, 3,
	                3, 2, 3, 2,
	                2, 3, 2, 3), nrow=4, ncol=4, byrow=TRUE)
# False
grid7 <- matrix(c(2,3,4,
	                3,4,2,
	                4,2,3), nrow=3, ncol=3, byrow=TRUE)
# False
grid8 <- matrix(c(-1,-2,-3,
	                -2,-3,-1,
	                -3,-1,-2), nrow=3, ncol=3, byrow=TRUE)
# False
grid9 <- matrix(c(1, 1, 1,
  1, 1, 2,
  1, 2, 3), nrow=3, ncol=3, byrow=TRUE)

# True
grid10 <- matrix(c(1), nrow=1, ncol=1, byrow=TRUE)
```


```{r prototype}
grid1 <- matrix(c(1,2,3,
	                3,1,2,
	                2,3,1), nrow=3, ncol=3, byrow=TRUE)

grid1 <- as.data.frame(grid1)

series_sum <- sum(1:3)

grid1 %>% 
  as.data.frame(.) %>% 
	dplyr::summarise(across(1:3, sum))  %>%
	dplyr::mutate(across(1:3, ~ case_when(. == series_sum ~ 1,
																				. != series_sum ~ 0)))
```

## Function

```{r sub_suduko, echo=TRUE}

sub_suduko <- function(data, N) {

      data    <- as.data.frame(data) 
      df_rows <- dim(data)[1]
      df_cols <- dim(data)[2]
      
			series_sum <- sum(1:N)

			cols_check <-	data %>%
              				dplyr::summarise(across(1:df_cols, sum)) %>%
			  	            dplyr::summarise(across(1:df_cols, abs))  %>%  
              				dplyr::mutate(across(1:df_cols, ~ case_when(. == series_sum ~ 1,
              																										. != series_sum ~ 0))) %>% 
			                mutate(sum = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>% 
			                pull(sum) 

			rows_check <- data %>% 
                			t(.) %>%
			                as.data.frame(.) %>% 
                			dplyr::summarise(across(1:df_rows, sum)) %>%
			  	            dplyr::summarise(across(1:df_rows, abs))  %>%  
                			dplyr::mutate(across(1:df_rows, ~ case_when(. == series_sum ~ 1,
                																									. != series_sum ~ 0))) %>% 
			  			        mutate(sum = rowSums(across(where(is.numeric)), na.rm = TRUE)) %>% 
			                pull(sum) 
			
			rows     <- if(df_rows - rows_check == 0) 'yes' else 'no' 
			cols     <- if(df_rows - cols_check == 0) 'yes' else 'no'
			if(rows == 'yes' & cols == 'yes') 'yessir!' else 'no sir' 
			
}

grid1 <- matrix(c(1,2,3,
	                3,1,2,
	                2,3,1), nrow=3, ncol=3, byrow=TRUE)

grid2 <- matrix(c(1, 2, 3,
                  1, 2, 3,
                  1, 2, 3), nrow=3, ncol=3, byrow=TRUE)

grid3 <- matrix(c(1, 1, 1,
                   2, 2, 2,
                   3, 3, 3), nrow=3, ncol=3, byrow=TRUE)

grid4 <- matrix(c(1000, -1000, 6,
                  2, 3, 1,
                  3, 1, 2), nrow=3, ncol=3, byrow=TRUE)

grid5 <- matrix(c(0), nrow=1, ncol=1, byrow=TRUE)

grid6 <- matrix(c( 3, 2, 3, 2,
                   2, 3, 2, 3,
                   3, 2, 3, 2,
                   2, 3, 2, 3), nrow=4, ncol=4, byrow=TRUE) 

grid7 <- matrix(c( 2, 3, 4,
                   3, 4, 2,
                   4, 2, 3), nrow=3, ncol=3, byrow=TRUE)   

grid8 <- matrix(c( -1,-2,-3,
                   -2,-3,-1,
                   -3,-1,-2), nrow=3, ncol=3, byrow=TRUE) 

grid9 <- matrix(c(1,1,1,
                  1,1,2,
                  1,2,3), nrow=3, ncol=3, byrow=TRUE) 

grid10 <- matrix(c(1), nrow=1, ncol=1, byrow=TRUE)

grid11 <- matrix(c(1,-2,-3,
                   -2,-3,1,
                   -3,1,-2), nrow=3, ncol=3, byrow=TRUE) 

sub_suduko(grid11, 3) 

```

```{r negatives}

# it seems like more needs to be done for this to really work with negative values 

grid11 %>% 
  as.data.frame(.) %>% 
	dplyr::summarise(across(1:3, sum)) # %>%
	# dplyr::summarise(across(1:3, abs))  %>%  
	# dplyr::mutate(across(1:3, ~ case_when(. == series_sum ~ 1,
	# 																			. != series_sum ~ 0)))
```

